---
title: "Iris Classifier"
author: "A Kissinger"
date: "11/26/2020"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=30), tidy=TRUE)
```

## Iris Classifier

Use the slider bars to input both the sepal and the petal lengths and widths of your flower to find out to within a 93.65% accuracy what species your iris is; given it is one of the three species in the data set *Iris setosa, versicolor or virginica.* There is an rpart model that runs under the hod that is used to calculate the predicted species. 

## Slide with Bullets

- I am using an inline server and ui for this setup, but I have also split this off into server.R and ui.R files all of which can be found here: All code
- Bullet 2
- Bullet 3

## Iris Type Predictor

```{r, message=FALSE, warning=FALSE,suppressPackageStartupMessages}

library(caret)
library(shiny)
library(datasets)

data(iris)
set.seed(10)

inBuild <- createDataPartition(y=iris$Species, p=0.75, list=FALSE)
iris.test <- iris[-inBuild,]
iris.train <- iris[inBuild,]

inBuild <- createDataPartition(y=iris.train$Species, p=0.75, list=FALSE)
iris.valid <- iris[-inBuild,]
iris.train <- iris[inBuild,]

shinyApp(
        ui=fluidPage(
  
    fluidRow(
    # Sidebar with a slider input for number of bins
    column(5, 
        wellPanel(
            sliderInput("SepalLength", "How long is the sepal?", 4.3, 7.3, value=5.8),
            sliderInput("SepalWidth", "How wide in the sepal?", 2.0, 4.4, value=3.2 ),
            sliderInput("PetalLength", "How long are the petals?", 1.0, 6.3, value=3.65),
            sliderInput("PetalWidth", "How wide are the petals?", 0.1, 2.5, value=1.3)
        )),

        # Show a plot of the generated distribution
        column(5,
            h3("With these inputs your iris has a 93.65% chance to be of species:"),
            h3(em(textOutput("pred1"))) 
    ))
),


server=function(input, output) {

    rpartout <- train(Species ~ .,method="rpart",data=iris.train, tuneLength=7)
    
    pred.rpart <- reactive({
        Sepal.Length <- input$SepalLength
        Sepal.Width <- input$SepalWidth
        Petal.Length <- input$PetalLength
        Petal.Width <- input$PetalWidth
        
        inputs <- as.data.frame(Sepal.Length)
        inputs <- cbind(inputs, Sepal.Width, Petal.Length, Petal.Width)
        as.character(predict(rpartout, newdata = inputs))
        # cmatrix.tree <-  confusionMatrix(pred.tree, iris.valid$Species)
        
    })
    
    
   output$pred1 <- renderText({
        pred.rpart()
    })
}
)
```

Slide With Plot
========================================================

```{r, echo=FALSE}
plot(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

