---
title: "Iris Classifier"
author: "A Kissinger"
date: "11/26/2020"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=30), tidy=TRUE)
```

## How to Use the Iris Classifier

Use the slider bars to input both the sepal and the petal lengths and widths of your flower to find out to within a 91.67% accuracy what species your iris is; given it is one of the three species in the data set *Iris setosa, versicolor or virginica.*


## Iris Type Predictor

```{r, message=FALSE, warning=FALSE,suppressPackageStartupMessages}

library(caret)
library(shiny)
library(datasets)

data(iris)
set.seed(1207)
inBuild <- createDataPartition(y=iris$Species, p=0.75, list=FALSE)
iris.test <- iris[-inBuild,]
iris.train <- iris[inBuild,]

 inBuild2 <- createDataPartition(y=iris.train$Species, p=0.75, list=FALSE)
 iris.valid <- iris.train[-inBuild2,]
 iris.train <- iris.train[inBuild2,]


shinyApp(
        ui=fluidPage(
  
    fluidRow(
    # Sidebar with a slider input for number of bins
    column(5, 
        wellPanel(
            sliderInput("SepalLength", "How long is the sepal?", 4.3, 7.3, value=5.8),
            sliderInput("SepalWidth", "How wide in the sepal?", 2.0, 4.4, value=3.2 ),
            sliderInput("PetalLength", "How long are the petals?", 1.0, 6.3, value=3.65),
            sliderInput("PetalWidth", "How wide are the petals?", 0.1, 2.5, value=1.3)
        )),

        # Show a plot of the generated distribution
        column(5,
            h3("With these inputs your iris has a 91.67% chance to be of species:"),
            h3(em(textOutput("pred1"))) 
    ))
),


server=function(input, output) {

    rpartout <- train(Species ~ .,method="rpart",data=iris.train, tuneLength=7)
    
    pred.rpart <- reactive({
        Sepal.Length <- input$SepalLength
        Sepal.Width <- input$SepalWidth
        Petal.Length <- input$PetalLength
        Petal.Width <- input$PetalWidth
        
        inputs <- as.data.frame(Sepal.Length)
        inputs <- cbind(inputs, Sepal.Width, Petal.Length, Petal.Width)
        as.character(predict(rpartout, newdata = inputs))
        # cmatrix.tree <-  confusionMatrix(pred.tree, iris.valid$Species)
        
    })
    
    
   output$pred1 <- renderText({
        pred.rpart()
    })
}
)
```


## Further Model Information

There is an rpart model running under the hood that is used to calculate the predicted species.

The dataset that was used to train the data was Edgar Anderson's Iris Data that comes with base R. The datset includes 150 cases of which I split into 75% and 25% training and testing sets respectively. The training data was then split again into 75% and 25% training and validation sets respectively. The final number of cases per dataset was: training=87, testing=36, validation=27. I tried both random forest and rpart models and the rpart won with a 96.3% accuracy on the validation set. The overall accuracy of the testing set using the rpart model was 91.67%. 

## Code Files:

I am using an inline server and ui for this setup, but I have also split these into server.R and ui.R files that can be found below: 
["Click this link"](https://github.com/apriljkissinger/datasciencecourserajohnhopkins/tree/master/IrisClassifier)

